% !TeX root = ../sustechthesis-example.tex

\begin{abstract}
  随着大语言模型（LLMs）在云计算环境中的广泛部署，其推理服务面临三类
  核心挑战：其一，现有批处理策略粗放且部署配置忽视网络拓扑异构性，导致
  GPU利用率仅为20\%--40\%并产生严重SLO违约；其二，预填充（Prefill）与解
  码（Decode）阶段资源需求天然互补（计算密集vs内存密集），但现有PD分离
  系统配置粒度过粗，互补性资源大量闲置；其三，调度决策与KV Cache状态紧
  耦合，前缀缓存感知路由引发持续性负载倾斜，系统难以适应动态异构负载。
  据统计，云平台中约90\%的人工智能计算资源用于模型推理而非训练，而现有
  系统受限于粗粒度的静态资源配置，在突发流量下因扩容滞后导致严重的SLO
  违约。针对这三类挑战，本文围绕\textbf{面向大模型的推理任务批式调度和
  KV缓存优化}这一主题，构建了从静态优化到动态调控的技术体系，实现了对
  LLM推理服务资源管理的系统性优化。

  首先，针对\textbf{挑战①}（批处理策略粗放、部署配置忽视网络拓扑异构
  性），本文提出了UELLM（Unified and Efficient LLM Inference Serving）统
  一高效批式调度与部署框架。该框架通过微调ChatGLM3-6B模型对请求输出长度
  进行预测，分桶准确率达到99.51\%，为后续批处理优化提供关键先验知识。在
  此基础上，设计了SLO与输出长度驱动的动态批处理算法（SLO-ODBS），通过权
  重化目标函数平衡总延迟与总输出长度，采用贪心策略将长度相近的请求组合
  成批，有效减少KV Cache冗余填充和计算浪费。同时，提出了基于动态规划的
  高效低延迟资源分配算法（HELR），综合考虑集群网络拓扑异构性
  （NVLink/PCIe带宽差异）和GPU计算能力差异，自动优化层到GPU的映射策略，
  支持高利用率（HE）与低延迟（LR）两种模式灵活切换。在4-GPU异构集群上的
  验证表明，UELLM相比Morphling和S$^3$等先进方案，推理延迟降低72.3\%至
  90.3\%，GPU利用率提升1.2倍至4.1倍，吞吐量提高1.92倍至4.98倍，并实现
  零SLO违约。

  其次，针对\textbf{挑战②}（PD分离架构资源互补性未被有效利用）与
  \textbf{挑战③}（调度决策与KV Cache状态紧耦合、难以适应动态异构负载），
  本文提出了BanaServe细粒度弹性资源管理与KV缓存优化框架。该框架创新性地
  引入了模块级细粒度迁移机制，包括层级（Layer-level）权重迁移与注意力级
  （Attention-level）KV Cache迁移：前者支持连续Transformer层的动态重配置，
  实现粗粒度负载均衡，有效应对挑战②中计算--显存资源失衡的问题；后者通
  过按注意力头维度切分KV Cache，选择性将部分注意力头状态迁移至辅助GPU，
  实现细粒度计算卸载，且无需传输模型权重，迁移开销极低，进一步充分利用
  Prefill与Decode阶段的互补性资源（对应挑战②）。针对挑战③，框架设计了
  全局KV Cache存储与层间流水线重叠传输机制，通过解耦缓存状态与计算位置
  消除前缀缓存对调度决策的约束，使路由器可基于纯负载指标进行调度，同时
  利用Transformer逐层计算特性将第$i$层计算与第$i+1$层KV Cache预取重叠，
  隐藏通信延迟；在此基础上提出了基于实时负载感知的请求调度算法，周期性
  测量各Prefill实例的综合负载（计算+内存利用率），将新请求分发至负载最轻
  的实例，配合动态迁移机制实现快速负载均衡。在13B参数模型
  （LLaMA-13B/OPT-13B）和公开基准（Alpaca/LongBench）上的评估表明，
  BanaServe相比vLLM吞吐量提升1.2倍至3.9倍、总处理时间降低3.9\%至78.4\%，
  相比DistServe吞吐量提升1.1倍至2.8倍、延迟降低1.4\%至70.1\%，并在突发
  流量场景下展现出优异的鲁棒性。

  本文所提出的方法在真实GPU集群和公开基准数据集上得到了充分验证，构建的
  “静态批处理优化--动态层级资源调控”技术体系为大模型推理服务的高效部署
  与资源管理提供了新的理论依据和技术路径，对推动大模型技术在智能客服、
  内容创作、知识检索等延迟敏感场景的规模化应用具有重要意义。

  \thusetup{
    keywords = {大语言模型推理, 批式调度, PD分离架构, 动态批处理,
      异构部署优化},
  }
\end{abstract}

\begin{abstract*}
  With the widespread deployment of Large Language Models (LLMs) in cloud
  computing environments, inference services face three core challenges:
  first, coarse-grained batching strategies and deployment configurations
  that neglect network topology heterogeneity result in GPU utilization of
  only 20\%--40\% and severe SLO violations; second, the naturally
  complementary resource demands of the prefill (compute-intensive) and
  decode (memory-intensive) phases remain unexploited due to coarse-grained
  configurations in existing PD-disaggregated systems; third, scheduling
  decisions are tightly coupled with KV Cache state, causing persistent load
  skew from prefix cache-aware routing and rendering systems unable to adapt
  to dynamic heterogeneous workloads. Approximately 90\% of AI computing
  resources in cloud platforms are dedicated to model inference rather than
  training, yet existing systems suffer from coarse-grained static resource
  configurations and severe SLO violations due to delayed scaling during
  bursty traffic. To address these three challenges, this thesis focuses on
  ``Batch Scheduling and KV Cache Optimization for Large Language Model
  Inference,'' constructing a technical architecture spanning from static
  optimization to dynamic scaling for systematic LLM inference resource
  management.

  Firstly, to address \textbf{Challenge~1} (coarse-grained batching
  strategies and deployment configurations that neglect network topology
  heterogeneity), this thesis proposes UELLM (Unified and Efficient LLM
  Inference Serving), a unified batch scheduling and deployment optimization
  framework. The framework employs a fine-tuned ChatGLM3-6B model to predict
  request output lengths, achieving a bucketing accuracy of 99.51\% to
  provide critical prior knowledge for subsequent batching optimization. The
  SLO and Output-Driven Dynamic Batch Scheduler (SLO-ODBS) balances total
  latency and total output length through a weighted objective function and
  adopts a greedy strategy to group requests with similar lengths into
  batches, effectively reducing KV Cache redundancy and computational waste.
  Furthermore, the High-Efficiency Low-Latency Resource Allocation (HELR)
  algorithm based on dynamic programming automatically optimizes
  layer-to-GPU mapping by considering cluster network topology heterogeneity
  (NVLink/PCIe bandwidth variations) and GPU computational capability
  differences, supporting flexible switching between High Efficiency (HE)
  and Low Latency (LR) modes. Validation on a 4-GPU heterogeneous cluster
  demonstrates that UELLM reduces inference latency by 72.3\%--90.3\%,
  improves GPU utilization by 1.2$\times$--4.1$\times$, and increases
  throughput by 1.92$\times$--4.98$\times$ compared to state-of-the-art
  methods such as Morphling and S$^3$, while achieving zero SLO violations.

  Secondly, targeting \textbf{Challenge~2} (unexploited complementary
  resource demands between Prefill and Decode phases in PD-disaggregated
  architectures) and \textbf{Challenge~3} (tight coupling between scheduling
  decisions and KV Cache state, preventing adaptation to dynamic
  heterogeneous workloads), this thesis presents BanaServe, a fine-grained
  elastic resource management and KV cache optimization framework. The
  framework introduces module-level migration mechanisms, including
  layer-level weight migration and attention-level KV Cache migration: the
  former supports dynamic reconfiguration of consecutive Transformer layers
  for coarse-grained load rebalancing, effectively addressing the
  compute-memory imbalance in Challenge~2; the latter achieves fine-grained
  computation offloading by partitioning KV Cache along the attention head
  dimension and selectively migrating partial attention head states to
  auxiliary GPUs without transferring model weights, further exploiting the
  complementary resources between Prefill and Decode phases (Challenge~2).
  To address Challenge~3, a Global KV Cache Store with layer-wise pipelined
  transmission is designed to eliminate constraints imposed by prefix caching
  on scheduling decisions by decoupling cache state from compute placement,
  enabling purely load-based routing while hiding communication latency by
  overlapping the computation of layer $i$ with the prefetching of KV Cache
  for layer $i+1$. A real-time load-aware request scheduling algorithm is
  further proposed, which periodically measures the combined load (compute
  and memory utilization) of each prefill instance and dispatches new
  requests to the least-loaded instance, working in coordination with dynamic
  migration for rapid load balancing. Evaluations on 13B-parameter models
  (LLaMA-13B/OPT-13B) and public benchmarks (Alpaca/LongBench) demonstrate
  that BanaServe achieves 1.2$\times$--3.9$\times$ higher throughput and
  3.9\%--78.4\% lower total processing time compared to vLLM, and
  1.1$\times$--2.8$\times$ higher throughput with 1.4\%--70.1\% lower
  latency compared to DistServe, while exhibiting superior robustness under
  bursty traffic scenarios.

  The proposed methods have been extensively validated on real GPU clusters
  and public benchmarks. The technical architecture of ``Static Batching
  Optimization~-- Dynamic Hierarchical Resource Scaling'' provides novel
  theoretical foundations and technical pathways for efficient deployment
  and resource management of LLM inference services, and holds significant
  importance for promoting the large-scale application of LLM technologies
  in latency-sensitive scenarios such as intelligent customer service,
  content creation, and knowledge retrieval.

  \thusetup{
    keywords* = {Large Language Model Inference, Batch Scheduling,
      PD Disaggregated Architecture, Dynamic Batching,
      Heterogeneous Deployment Optimization},
  }
\end{abstract*}
